---
title: 'Reproducible Research: Peer Assignment 1'
output: html_document
---

#Loading and preprocessing the data
Read in the data and process, removing incomplete rows.

```{r echo=TRUE}
myData <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
myData <- na.exclude(myData)
rownames(myData) <- 1:nrow(myData)
```

#What is mean total number of steps taken per day?
Construct a histogram of the average number of steps per day.
```{r echo=TRUE}
library(ggplot2)
ggplot(myData, aes(date,steps)) +
  geom_bar(stat = "identity", width = 1.0) +
  labs(title = "Histogram of Total Number of Steps Taken",
       x = "Date", y = "Steps")
```

Calculate the mean and median total number of steps per day.
```{r echo=TRUE}
totalSteps <- aggregate(myData$steps,list(myData$date),FUN="sum")[,2]
mean(totalSteps)
median(totalSteps)
```

#What is the average daily activity pattern?
Make a time series plot of the 5-minute interval, averaged across all days.

```{r echo=TRUE}
avgSteps <- aggregate(myData$steps,
                      list(as.numeric(as.character(myData$interval))),
                      FUN = "mean")
names(avgSteps) <- c("period","averageSteps")
ggplot(avgSteps, aes(period, averageSteps)) +
        geom_line(size = 1.0) +
        labs(title = "Time Series Plot of 5 Min Interval",
             x = "5 Min interval", y = "Average Steps Taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains
the maximum number of steps?
```{r echo=TRUE}
avgSteps[avgSteps$averageSteps == max(avgSteps$averageSteps),]
```

#Imputing missing values
Calculate and report the total number of missing values in the dataset.
```{r echo=TRUE}
originalData <- read.csv("activity.csv",
                         colClasses = c("integer", "Date", "factor"))
sum(is.na(originalData))
```

Use the impute function from the Hmisc package to impute average number of steps
across all days.
```{r echo=TRUE}
library(Hmisc)
imputed <- originalData
imputed$steps <- with(imputed, impute(steps, mean))
```

Create the histogram for the imputed dataset.
```{r echo=TRUE}
ggplot(imputed, aes(date,steps)) +
        geom_bar(stat = "identity", width = 1.0) +
        labs(title = "Histogram of Total Number of Steps Taken",
             x = "Date", y = "Steps")
```

Look at the averages and medians of the original vs imputed datasets:
```{r echo=TRUE}
totalSteps.imputed <- aggregate(imputed$steps,list(imputed$date),FUN="sum")[,2]
mean(totalSteps.imputed)
mean(totalSteps)
median(totalSteps.imputed)
median(totalSteps)
```

#Are there differences in activity patterns between weekdays and weekends?
Create and call a function that uses the weekday() function to assign whether
the date is a weekday or weekend day.
```{r echo=TRUE}
dayOfWeek <- function(date){
        if(weekdays(date) %in% c("Saturday","Sunday")){
                return("weekend")
        }
        else{
                return("weekday")
        }
}

day <- sapply(imputed$date, dayOfWeek)
imputed$day <- as.factor(day)
```

Calculate the average steps per interval, grouped by weekend vs weekday.
```{r echo=TRUE}
avgSteps.imputed <- aggregate(imputed$steps,
                              list(as.numeric(as.character(imputed$interval)),
                                   imputed$day),
                              FUN = "mean")
names(avgSteps.imputed) <- c("period","dayType","averageSteps")
```

Plot the average steps per interval, grouped by weekend vs weekday.
```{r echo=TRUE}
ggplot(avgSteps.imputed, aes(period, averageSteps)) +
        geom_line(size = 1.0) +
        facet_grid( . ~ dayType) +
        labs(title = "Time Series Plot of 5 Min Interval",
             x = "5 Min interval", y = "Average Steps Taken")
```